---
title: 'Cognitive Aspects of Software Development: a SMS'
output:
  html_document: 
    df_print: tibble
    fig_caption: yes
    code_folding: hide
  pdf_document: default
editor_options:
  chunk_output_type: inline
---

```{r}
# Keep things nice and tidy, all libraries go here
library(readxl)
library(tidyverse)
library(knitr)
library(kableExtra)
library(svglite)
library(plotly)
library(scales)
library(ggsci)
library(RColorBrewer)
# for color palettes
library(paletti)
library(pals)
```
```{r}
data <- read_excel("data/data_post_IEEE_fix.xlsx", skip = 1, sheet = "Extraction and classification (")
data <- data %>% filter(is.na(Exclude))
# Years without any publication (for easy slicing)
years_no_publications <- c("1974",
                           "1975",
                           "1976",
                           "1978")

# LABELS so slicing will not become a mess
swebok_areas_labels = c("SR",
                        "SD",
                        "SC",
                        "ST",
                        "SM",
                        "SCM",
                        "SEM",
                        "SEP",
                        "SEMM",
                        "SQ",
                        "SEPP",
                        "SEE",
                        "CF",
                        "MF",
                        "EF")

swebok_areas_labels_no_foundation = c("SR",
                                      "SD",
                                      "SC",
                                      "ST",
                                      "SM",
                                      "SCM",
                                      "SEM",
                                      "SEP",
                                      "SEMM",
                                      "SQ",
                                      "SEPP",
                                      "SEE")

swebok_areas_labels_long = c("Requirements",
                                "Design",
                                "Construction",
                                "Testing",
                                "Maintainance",
                                "Config. Mgmt.",
                                "SE Mgmt.",
                                "SE Processes",
                                "SE Models&Methods",
                                "Software Quality",
                                "SE Prof. Practice",
                                "SE Economics")

swebok_areas_labels_long_all = c("Computing Found.",
                                 "Engineering Found.",
                                 "Mathematical Found.",
                                "Construction",
                                "Config. Mgmt.",
                                "Design",
                                "SE Economics",
                                "SE Mgmt.",
                                "SE Models&Methods",
                                "SE Processes",
                                "SE Prof. Practice",
                                "Maintainance",
                                "Software Quality",
                                "Requirements",
                                "Testing")

cognitive_concepts_labels <- c("Attention",
                               "Selective attention",
                               "Divided attention",
                               "Sustained attention",
                               "Memory",
                               "Working memory",
                               "Short-term memory",
                               "Long-term memory",
                               "Cognitive load",
                               # "Cognitive control",
                               "Intrinsic CL",
                               "Extrinsic CL",
                               "Perception",
                               "Problem solving",
                               "Reasoning",
                               "Decision making",
                               "Cognitive biases",
                               "Knowledge",
                               "Explicit knowledge",
                               "Tacit knowledge",
                               "Techn. tacit knowl.",
                               "Cogn. tacit knowl.")

measures_labels <- c("Qualit. measures",
                      "Fieldwork",
                      "Interview",
                      "Task-based",
                      "Open observation",
                      "Quantit. measures",
                      "Task performance",
                      "Physiological meas.",
                      "Subjective ratings",
                      "Behavioral meas.")

paper_types <- c("Solution Proposal (example)", 
                 "Empirical Study", 
                 "Experience Papers",
                 "Philosophical Papers",
                 "Opinion Papers")


high_level_concepts <- c("Attention", 
                         "Memory",
                         "Cognitive load",
                         "Perception",
                         "Reasoning",
                         "Cognitive biases",
                         "Knowledge")
# COLORS
tol9qualitative <- c("#332288",
                  "#88CCEE",
                  "#44AA99",
                  "#117733",
                  "#999933",
                  "#DDCC77",
                  "#CC6677",
                  "#882255",
                  "#AA4499")

NPG_modified <- c("#F5E144",
               "#4DBBD5FF",
               "#00A087FF",
               "#3C5488FF",
               "#F39B7FFF",
               "#8491B4FF",
               "#91D1C2FF",
               "#DC0000FF",
               "#7E6148FF")
col25 <- c(
  "dodgerblue2", "#E31A1C",
  "green4",
  "#6A3D9A",
  "#FF7F00",
  "black", "gold1",
  "skyblue2", "#FB9A99",
  "palegreen2",
  "#CAB2D6",
  "#FDBF6F",
  "gray70", "khaki2",
  "maroon", "orchid1", "deeppink1", "blue1", "steelblue4",
  "darkturquoise", "green1", "yellow4", "yellow3",
  "darkorange4", "brown"
)
# Necessary for groupying by high-level category
add_high_level_concepts_to_data <- function(data) {
  data %>%
    mutate(Concept = case_when(
    Taxonomy %in% c("Attention", "Selective attention", "Divided attention", "Sustained attention") ~ "Attention",
    Taxonomy %in% c("Memory", "Working memory", "Short-term memory", "Long-term memory") ~ "Memory",
    Taxonomy %in% c("Cognitive control", "Cognitive load", "Extrinsic CL", "Intrinsic CL") ~ "Cognitive load",
    Taxonomy == "Perception" ~ "Perception",
    Taxonomy %in% c("Problem solving", "Reasoning", "Decision making") ~ "Reasoning",
    Taxonomy %in% c("Cognitive biases") ~ "Cognitive biases",
    Taxonomy %in% c("Knowledge", "Explicit knowledge", "Tacit knowledge",  "Techn. tacit knowl.", "Cogn. tacit knowl.") ~ "Knowledge",
    ))
  }
```

# Visualizing number of publications over time
```{r}
ggplot(data, aes(x=as.factor(Year))) +
  geom_bar() +
  ylab("Number of publications") +
  xlab("Year") +
  geom_text(stat='count', aes(label=..count..), vjust=2, color="white", size = 2.5) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
ggsave("PDFs/yearly_distribution.pdf", dpi = "screen")
```


```{r}
# Cleaning not needed values
data<-data %>%
    mutate(Academia = replace(Academia, Academia == "?", NA)) %>%
    mutate(Industry = replace(Industry, Industry == "?", NA))

data<-data %>%
  mutate(Type = case_when(is.na(Academia) & is.na(Industry) ~ "None",
                       Academia == "1" & is.na(Industry) ~ "Academia",
                       Industry == "1" & is.na(Academia) ~ "Industry",
                       TRUE ~ "Both"))
```

# Number of publications according to their type
```{r}
data %>%
  mutate(Type = fct_infreq(Type, ordered = T)) %>%
ggplot(aes(x=Type)) +
  geom_bar(width = .5) +
  xlab("Type of publication") +
  ylab("Number of publications") +
  geom_text(stat='count', aes(label=..count..), vjust=3, color="white", size = 4) +
  theme_bw()
ggsave("PDFs/academia_industry_distribution.pdf", dpi = "screen",  width = unit(10, "inch"), height = unit(6.5, "inch"))
```

# Number of publications categorized according to SWEBoK Areas.
A publication can be in more than one category at the same time.

```{r}
data %>%
  select(all_of(swebok_areas_labels)) %>% # selecting columns corresponding to the SWEBoK Areas
  mutate_all(replace_na,0) %>%
  mutate(SEPP = as.numeric(SEPP)) %>%
  summarise_all(sum) %>%
  gather(key = "SWEBOKArea", value = "publications", 1:15) %>%
  arrange(-publications) %>%
  mutate(SWEBOKArea = factor(SWEBOKArea, SWEBOKArea)) %>%
  ggplot(aes(x=SWEBOKArea, y=publications)) +
  geom_bar(stat="identity") +
  geom_text(aes(label=publications), vjust=-0.3, color="black", size = 4) +
  xlab("SWEBoK Area") +
  ylab("Number of publications") +
  theme_bw()

ggsave("PDFs/swebok_distribution.pdf", dpi = "screen",  width = unit(10, "inch"), height = unit(6.5, "inch"))
```

# Co-occurrences of SWEBoK Areas
```{r}
swebokareas<-data %>%
  select(all_of(swebok_areas_labels)) %>% # selecting columns corresponding to the SWEBoK Areas
  mutate_all(replace_na,0) %>%
  as.matrix() %>%
  crossprod()

swebokareas %>%
  kable()

plot_ly(x=swebok_areas_labels, y=swebok_areas_labels, z=swebokareas, type="heatmap")
```

```{r}
x <- data %>% select(all_of(swebok_areas_labels), all_of(cognitive_concepts_labels)) %>%
  mutate_all(replace_na, 0) %>%
  mutate(`Problem solving` = as.numeric(`Problem solving`)) %>%
  gather(key="SWEBOK", value = pubs, swebok_areas_labels) %>% # use SWEBOK area as factor
  filter(pubs > 0) %>% # select areas for which there are publications
  group_by(SWEBOK, .add=T) %>%
  # xtally() %>% # number of publication for each area
  # select(-pubs) %>%  # remove pubs to reuse it later
  gather(key = "Taxonomy", value = "count", cognitive_concepts_labels) %>%  # count publications in each cognitive taxonomy area
  mutate(label = str_replace(as.character(count), "^0", "")) # add label for later
```
```{r}
# Bubble plot
x <- arrange(x, Taxonomy)
xf<-x$Taxonomy
xfu<-unique(xf)
x$Taxonomy<-factor(xf,levels=xfu)

p<-ggplot(x)
p + geom_point(aes(x = fct_infreq(SWEBOK), y = fct_rev(Taxonomy), size=count), shape=21, fill="white", alpha=0.60) +
geom_text(aes(x = fct_infreq(SWEBOK), y = fct_rev(Taxonomy), label=label), size=2) +
theme(axis.text.x = element_text(angle = 45, hjust = 1.1, size=9,colour="black"), axis.text.y = element_text(size=8,colour="black"), axis.title.x = element_text(size=10), axis.title.y = element_text(size=10,colour = "black",vjust=0.12), panel.grid.major = element_line(linetype = "dashed", size=0.1, color="black"))+
  labs(x="SWEBOK Area",y = "Taxonomy Area") + theme_bw()
ggsave("PDFs/swebok_taxonomy_bubble.pdf",  dpi = "screen")
```
```{r}
# Preparing the dataset for analysing the research methods
data<-data %>%
  mutate(Quantitative = case_when(`Quantit. measures` == 1 | `Task performance` == 1 | `Physiological meas.` == 1 | `Subjective ratings` == 1 | `Behavioral meas.` == 1 ~ 1)) %>%
  mutate(Quantitative = replace_na(Quantitative, 0)) %>%
  mutate(Qualitative = case_when(Fieldwork == 1 | Interview == 1 | `Qualit. measures` == 1 | `Task-based` == 1 | `Open observation` == 1 ~ 1)) %>%
  mutate(Qualitative = replace_na(Qualitative, 0)) %>%
  mutate(Both = if_else(Qualitative == 1 & Quantitative == 1, 1, 0))
```


# The graphs below are prepared for IEEE Software Submission

Number of publications per year according to SWEBOK areas
```{r}
# Creating a temp dataset with missing publications years (i.e., year for which there was no publication)

colourCount = length(swebok_areas_labels_long_all)
getPalette = colorRampPalette(brewer.pal(9, "Set1"))

data %>%
  filter(is.na(Exclude)) %>%
  select(c(Year, SR:EF)) %>%
  gather("SWEBOK", "publications", 2:16) %>%
  mutate_all(replace_na, 0) %>%
  mutate(publications = as.numeric(publications)) %>%
  group_by(Year,SWEBOK)  %>%
  summarise(total=sum(publications)) %>%
  ggplot(aes(x=as.factor(Year), fill=SWEBOK, y=total)) +
  geom_bar(stat="sum") +
  xlab("Year") + ylab("Publications") + 
  scale_fill_manual(values = getPalette(colourCount), labels=swebok_areas_labels_long_all) +
  scale_y_continuous(breaks = seq(1,33,2)) +
  guides(size = F) + theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 6)) + 
  theme(legend.key.size = unit(0.4, 'cm'), #change legend key size
        legend.key.height = unit(0.3, 'cm'), #change legend key height
        legend.key.width = unit(0.5, 'cm'), #change legend key width
        legend.title = element_text(size=10), #change legend title font size
        legend.text = element_text(size=8),
        legend.position=c(.15,.7)) 
ggsave("PDFs/years_swebok.pdf", dpi = "screen",  width = unit(10, "inch"), height = unit(6.5, "inch"))
```

### Evolution of research methods over the years
```{r}
data <-  data %>% complete(Year=seq(1973,2016))

data <-  data %>%
  mutate(research_method = if_else(Both==1, "Mixed", if_else(Qualitative==1, "Qualitative", "Quantitative")))  %>%
  filter(!is.na(research_method))

data %>%  ggplot(aes(x=as.factor(Year), fill=research_method)) + geom_bar() +
  scale_fill_discrete(name="Research method", labels = c("Mixed", "Qualitative", "Quantitative", "")) +
  xlab("Year") + ylab("Publications") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 5))

ggsave("PDFs/years_researchmethods.pdf",  dpi = "screen",  width = unit(10, "inch"), height = unit(6.5, "inch"))
```


### Prevalence of research methods in the SWEBOK areas
```{r}
data.swebok.researchmethod <- data %>%
  select(swebok_areas_labels, research_method) %>%
  mutate_all(replace_na,0) %>%
  filter(research_method != 0) %>%
  mutate(SEPP = as.numeric(SEPP)) %>% # quickfix. For some reason SEPP column is not cast to numeric
  group_by(research_method) %>%
  summarise_at(vars(swebok_areas_labels), sum) %>%
  gather("SWEBOK", "Publications", swebok_areas_labels)

data.swebok.researchmethod %>%
  ggplot(aes(x=reorder(SWEBOK, Publications, function(x){sum(x)}), y=Publications, fill=research_method)) + geom_bar(stat = "identity") +
  coord_flip() + xlab("SWEBOK areas") + scale_fill_discrete(name = "Research method")

ggsave("PDFs/SWEBOK_researchmethods.pdf",  dpi = "screen",  width = unit(10, "inch"), height = unit(6.5, "inch"))
```
# Distribution of publications
```{r}
data %>%
  filter(!is.na(Identifier)) %>%
  select(Identifier, all_of(cognitive_concepts_labels), measures_labels) %>%
  gather(Taxonomy, value, all_of(cognitive_concepts_labels)) %>%
  filter(!is.na(value)) %>%
  select(-value) %>%
  gather(Method, value, measures_labels) %>%
  filter(!is.na(value)) %>%
  arrange(Identifier) %>%
  select(-Identifier, -value) %>%
  group_by(Taxonomy, Method) %>%
  tally(name = "Amount") %>%
  filter(Method!='Qualit. measures' & Method!='Quantit. measures') %>%
  ggplot(aes(x=fct_relevel(Method, measures_labels), y=Taxonomy, size=Amount)) +
  geom_point(aes(alpha=0.8)) +
  scale_size_continuous(range = c(3, 12)) +
  xlab("Cognitive Assessment Procedures") +
  ylab("Cognitive Concepts") +
  geom_vline(xintercept = 4.5, size=0.5,  color="darkgrey") +
  annotate(geom="text", x=6.5, y=0.7, label="Quantitative", size=3, alpha=0.9) +
  annotate(geom="text", x=2.5, y=0.7, label="Qualitative", size=3, alpha=0.9) +
  theme(legend.position = "", axis.text.x = element_text(angle = 30, hjust = 1, size = 8))

ggsave("PDFs/taxonomy_methods.pdf",  dpi = "screen",  width = unit(10, "inch"), height = unit(6.5, "inch"))
```

```{r}
data %>%
  select(all_of(swebok_areas_labels_no_foundation), all_of(cognitive_concepts_labels)) %>%
  mutate_all(replace_na,0) %>%
  gather(Taxonomy, value2, cognitive_concepts_labels) %>%
  add_high_level_concepts_to_data() %>%
  gather(SWEBOK, value, all_of(swebok_areas_labels_no_foundation)) %>%
  count(SWEBOK, Concept, value, value2) %>%
  mutate(freq=ifelse(value==1 & value2==1, n, 0)) %>%
  distinct(SWEBOK, Concept, freq) %>%
  group_by(SWEBOK, Concept) %>%
  summarize(total=sum(freq)) %>%
  filter(total > 0) %>% 
  ungroup() %>%
  ggplot(aes(fct_relevel(SWEBOK, swebok_areas_labels_no_foundation), fct_rev(Concept), fill=total)) +
  geom_tile() + scale_fill_continuous(low="#fff9f7", high="red") +
  xlab("SWEBOK areas") + ylab("Cognitive Concepts") + guides(fill=guide_legend(title="Publications")) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 30, hjust = 1, size = 8)) +
  scale_x_discrete(labels = swebok_areas_labels_long)

ggsave("PDFs/taxomony_swebok_cooccurences.pdf",  width = unit(10, "inch"), height = unit(6.5, "inch"))
```



```{r}
data %>%
  select(cognitive_concepts_labels,  measures_labels) %>%
  mutate_all(replace_na,0) %>%
  gather(Taxonomy, value, cognitive_concepts_labels) %>%
  add_high_level_concepts_to_data() %>%
  gather(Method, value2, measures_labels) %>%
  count(Concept, Method, value, value2) %>%
  mutate(freq=ifelse(value==1 & value2==1, n, 0)) %>%
  mutate(freq=as.integer(freq)) %>%
  mutate(Method=fct_relevel(Method, measures_labels)) %>%
  filter(Method != 'Qualit. measures' & Method != 'Quantit. measures') %>%
  filter(freq > 0) %>%
  ggplot(aes(Method, Concept, fill=freq)) +
  geom_tile() +
  geom_vline(xintercept = 4.5, size=0.5,  color="darkgrey") +
  xlab("Cognitive Assessment Procedures") + ylab("Cognitive Concepts") +
  guides(fill=guide_legend(title="Publications")) +  # scale_x_discrete(labels=c("Fieldwork", "Interview", "Task-based", "Open observation", "Others", "Task performance", "Physiological meas.", "Subjective ratings", "Behavioral meas.", "Others")) + # not using measure_lables here since we need a catch-all "Others" category
   scale_fill_continuous(low="#fff9f7", high="darkgreen") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 30, hjust = 1, size = 8)) +
  annotate(geom="text", x=6.5, y=0.7, label="Quantitative", size=3, alpha=0.6) +
  annotate(geom="text", x=2.5, y=0.7, label="Qualitative", size=3, alpha=0.6) +

ggsave("PDFs/taxonomy_method_cooccurences.pdf",  dpi = "screen",  width = unit(10, "inch"), height = unit(6.5, "inch"))
```

```{r}
data %>%
  select(Year, cognitive_concepts_labels)%>%
  filter(Year > 0) %>%
  gather("Taxonomy", "publications", cognitive_concepts_labels) %>%
  mutate_all(replace_na,0) %>%
  mutate(publications=as.integer(publications)) %>%
  group_by(Year, Taxonomy) %>%
  summarise(total=sum(publications)) %>%
  ggplot(aes(as.factor(Year), total, fill=Taxonomy)) + geom_bar(stat="sum") +  xlab("Year") + ylab("Publications") +
  scale_fill_discrete(name = "Taxonomy Areas") + guides(size = F) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 6))
```

```{r}
cols_fill_taxonomy_years  <- get_scale_fill(get_pal(pals::stepped()))

 df.taxonomy  <- data %>%
  select(Year, all_of(cognitive_concepts_labels)) %>%
  gather("Taxonomy", "publications", cognitive_concepts_labels) %>%
  mutate_all(replace_na,0) %>%
  mutate(publications=as.integer(publications)) %>% # for some reseason recognized as char
  filter(publications>0)

# need to create a separated df to hold the percentage of publications within each year
data.percentage <-  df.taxonomy  %>%
  group_by(Year) %>%
  count(Taxonomy) %>%
  mutate(ratio = scales::percent(n/sum(n)))

df.taxonomy %>%
  ggplot(aes(x = as.factor(Year), fill = as.factor(Taxonomy))) +
  geom_bar(position="fill") +
  geom_text(data = data.percentage, aes(y = n,label = ratio), position = position_fill(vjust = 0.5), colour = "black", size = 1.3, alpha=0.5) +
  xlab("Year") + ylab("Publications %") +
  guides(size = F, fill=guide_legend(title="Concepts (including sub-concepts)")) +
  scale_y_continuous(labels = percent) +
  theme_minimal() + theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8)) +
  theme(legend.key.size = unit(.2, "cm"), legend.key.width = unit(0.2,"cm"), legend.title = element_text(size = 8), legend.text = element_text(size = 6)) +
cols_fill_taxonomy_years()

ggsave("PDFs/taxonomy_years.pdf", width = unit(10, "inch"), height = unit(6.5, "inch"))
```

```{r}
df.concepts <- df.taxonomy %>%
  add_high_level_concepts_to_data()
df.years <- data %>% filter(!(Year %in% years_no_publications)) %>% count(Year) # years without publications

ggplot() +
  geom_bar(data=df.concepts, aes(x=as.factor(Year), fill=Concept), position="fill") +
  geom_line(data=df.years, aes(x=as.factor(Year), y=n/max(n), group=1), size=0.8) +
  geom_point(data=df.years, aes(x=as.factor(Year), y=n/max(n), group=1)) +
  scale_y_continuous(labels = function(x)x*100, name="Publication %", sec.axis = sec_axis(name="Total publications", ~. * max(df.years$n), breaks=scales::breaks_extended(10))) +
  xlab("Year")  +
  theme(panel.background = element_blank(), axis.ticks.x = element_blank(), axis.title.x = element_text(margin = margin(-15,0,0,0, "pt")), axis.text.x = element_text(angle = 45, hjust = 1, size = 8, vjust = 2.4)) +
  scale_fill_manual(values = NPG_modified)

ggsave("PDFs/cc_cm_classification_wide_colored.pdf", width = unit(10, "inch"), height = unit(6.5, "inch"))
```
```{r}
data %>%
  filter(!is.na(publication_type)) %>%
  ggplot(aes(x=fct_rev(fct_infreq(publication_type)))) +
  geom_bar(stat="count", width=.4) +
  geom_text(stat='count', aes(label=..count..), hjust=-0.3, color="black", size = 3) +
  scale_x_discrete("Publication outlets", labels=c("Magazines", "Books",  "Workshops", "Conferences", "Journals" )) +
  scale_y_continuous("Number of publications", breaks = scales::breaks_extended(10)) +
  coord_flip() +
  theme_minimal() +
  theme(axis.text.y = element_text(angle = 90, hjust=.5, vjust=-0.3)) 
  
ggsave("PDFs/outlets.pdf",  dpi = "screen",  width = unit(10, "inch"), height = unit(6.5, "inch"))

```
```{r}
data %>%
  filter(!is.na(publication_type)) %>%
  select(all_of(cognitive_concepts_labels), publication_type) %>%
  mutate_all(replace_na,0) %>%
  gather(Taxonomy, value2, cognitive_concepts_labels) %>%
  add_high_level_concepts_to_data() %>%
  mutate(value2 = as.integer(value2)) %>%
  filter(value2 > 0) %>%
  ggplot(aes(x=fct_rev(fct_infreq(Concept)))) +
  geom_bar(stat="count", aes(fill=publication_type)) +
  xlab("Cognitive Concepts") +
  scale_y_continuous("Number of publications", breaks = scales::breaks_extended(10)) +
  coord_flip() +
  # scale_fill_discrete(name = "Publication outlets", labels = c("Books", "Conferences", "Journals", "Workshops", "Magazines")) +
   scale_fill_brewer(palette = "Dark2", labels=c("Books", "Conferences", "Journals", "Workshops", "Magazines")) + 
  scale_y_continuous(breaks = seq(0,170,10)) +
  theme_minimal() + 
  theme(legend.key.size = unit(0.7, 'cm'), #change legend key size
        legend.key.height = unit(0.6, 'cm'), #change legend key height
        legend.key.width = unit(0.8, 'cm'), #change legend key width
        legend.title = element_text(size=13), #change legend title font size
        legend.text = element_text(size=11),
        legend.position=c(.8,.2)) +
  labs(fill = "Publication Type") + 
  ylab(label = "Number of publications")


ggsave("PDFs/outlets_concepts.pdf",  dpi = "screen",  width = unit(10, "inch"), height = unit(6.5, "inch"))


```


```{r}
data %>%
  group_by(publication_type) %>%
  tally() %>%
  arrange(publication_type, n) %>%
  top_n(5) %>%
  kable(col.names = c("Outlet", "#"))
```

```{r}
data %>% 
  select(paper_types, high_level_concepts) %>% 
  mutate_all(replace_na, 0) %>% 
  gather(Taxonomy, value2, high_level_concepts) %>% 
  gather(Type, value, all_of(paper_types)) %>% 
  count(Type, Taxonomy, value, value2) %>% 
  mutate(freq=ifelse(value==1 & value2==1, n, 0)) %>% 
  distinct(Type, Taxonomy, freq) %>% 
  group_by(Type, Taxonomy)  %>% 
  summarize(total=sum(freq)) %>% 
  filter(total > 0) %>% 
  ungroup() %>% 
  ggplot(aes(reorder(Taxonomy, total, function(x){sum(x)}), total, fill=Type)) + 
  geom_bar(stat="identity") +
  coord_flip() + 
  xlab("Cognitive Concepts") + 
  ylab("Number of publications") + 
  scale_fill_brewer(palette = "Dark2") +
#  scale_fill_manual(values=c("#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00")) +
  theme_minimal() +
  theme(legend.key.size = unit(0.7, 'cm'), #change legend key size
        legend.key.height = unit(0.6, 'cm'), #change legend key height
        legend.key.width = unit(0.8, 'cm'), #change legend key width
        legend.title = element_text(size=13), #change legend title font size
        legend.text = element_text(size=11),
        legend.position=c(.8,.2)) 

ggsave("PDFs/taxonomy_paper_type.pdf",  dpi = "screen",  width = unit(10, "inch"), height = unit(6.5, "inch"))
```

